### 2017-07-02 related-work.tex
edits

\vspace*{-8pt}
\section{Related Work}
\label{sec:relatedwork}

\paragraph{Specifying weak isolation.}
Adya~\cite{adyaphd} specifies several weak isolation levels in terms
of \emph{dependency graphs} between transactions, and the kinds of
dependencies that are forbidden in each case. The operational nature
of Adya's specifications make them suitable for runtime monitoring and
anomaly detection~\cite{kemmevldb,feketesigmod08,pssi2011}, whereas
the declarative nature of our specifications make them suitable for
formal reasoning about program behaviour. Sivaramakrishnan \emph{et
al.}~\cite{pldi15} specify isolation levels declaratively as trace
well-formedness conditions, but their specifications implicitly assume
a complete trace with only committed transactions, making it difficult
to reason about a program as it builds the trace. Cerone \emph{et
al.}~\cite{gotsmanconcur15} specify isolation levels with atomic
visibility, but their specifications are also for complete traces.
Like ours, all the aforementioned specification frameworks use the
vocabulary introduced in~\cite{burckhardt14}. However, none of them
are equipped with a reasoning framework that can use such
specifications to verify programs under weak isolation.

>>>>

\vspace*{-6pt}
\section{Related Work}
\label{sec:relatedwork}

\paragraph{Specifying weak isolation.}

Adya~\cite{adyaphd} specifies several weak isolation levels in terms
of \emph{dependency graphs} between transactions, and the kinds of
dependencies that are forbidden in each case. The operational nature
of Adya's specifications make them suitable for runtime monitoring and
anomaly detection~\cite{kemmevldb,feketesigmod08,pssi2011}, whereas
the declarative nature of our specifications make them suitable for
formal reasoning about program behaviour. Sivaramakrishnan \emph{et
al.}~\cite{pldi15} specify isolation levels declaratively as trace
well-formedness conditions, but their specifications implicitly assume
a complete trace with only committed transactions, making it difficult
to reason about a program as it builds the trace. Cerone \emph{et
al.}~\cite{gotsmanconcur15} specify isolation levels with atomic
visibility, but their specifications are also for complete traces.
Like ours, all the aforementioned specification frameworks use the
vocabulary introduced in~\cite{burckhardt14}. However, none of them
are equipped with a reasoning framework that can use such
specifications to verify programs under weak isolation.

Recent work by Crooks \emph{et al.}~\cite{CPA+17} also explores the
use of a state-based interpretation of isolation as we do, and share
our position that enables specification of weak isolation that are not
tied to implementation-specific artifacts.  However, they do not
consider verification (manual or automated) of client programs, and it
is not immediately apparent if their specification formalism is
amenable for use within a verification toolchain.
